---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
#loading libraries
library(tidyverse)
library(plotly)
library(readr)
library(lubridate)
library(stringr)
library(ggplot2)
library(ggmap)
library(purrr)
library(gridExtra)

require("knitr")
opts_knit$set(root.dir = "C:/Users/rranadive/Downloads/nypd")
```


```{r }
#nypd-accidents data exported from Goodle BigQuery API
data <- read_csv("data/data_mydata2.csv")

#adding year
data$year <- year(data$timestamp)
data$month <- month(data$timestamp)
data$day <- day(data$timestamp)
data$hour <- hour(data$timestamp)
```

```{r}
#missing data
NAcol <- which(colSums(is.na(data)) > 0)
sort(colSums(sapply(data[NAcol], is.na)), decreasing = TRUE)
```

```{r}
data$borough <- as.factor(data$borough)
```

```{r}
table(data$borough,useNA = "always")
```

### Frequency distribution of Borough 

```{r}
options(scipen=10000)
ggplot(data=subset(data, !is.na(data$borough)), aes(x=borough))+geom_histogram(stat = "count",color="blue", fill="blue")+scale_y_continuous(breaks=c(100000,200000,300000),labels=c('100K','200K','300K'))+xlab("Major Cities of NY")+ylab("Number of Accidents")+theme_light()
```

```{r}
#registering Google Static Map API key
ggmap::register_google(key="AIzaSyBl1ERJg4Q6UUwsrdEPT7-iprbS9Zcm2z0")
```

```{r}
## getting New York centered map
p <- ggmap(get_googlemap(center = c(lon=-74.0060,lat=40.7128),
                    zoom = 10, scale =2,
                    maptype ='terrain'), color='color')
```

```{r}
p + geom_point(data = subset(data, !is.na(data$latitude)& !is.na(data$longitude)),na.rm = T, aes(x = longitude, y = latitude,color=borough)) + geom_jitter()
```

```{r}
ggsave("Nyaccidents.png", dpi=300)
```


```{r}
##latitude and longitude cannot be zero for NY:
## so replacing those 
data$latitude[data$latitude==0] <- NA
data$longitude[data$longitude==0] <- NA
```

```{r}
p + geom_point(data = subset(data, (!is.na(data$latitude) & !is.na(data$longitude) & !is.na(data$borough))),na.rm = T, aes(x = longitude, y = latitude,color=borough)) + geom_jitter()
```

## Getting data only for Brooklyn:
```{r}
databrooklyn <- filter(data , borough=="BROOKLYN")
```

## Brooklyn Centered Map:
```{r}
brooklyn <- ggmap(get_googlemap(center = c(lon=-73.95,lat=40.655),
                    zoom = 12, scale =2,
                    maptype ='terrain'), color='color')
```

```{r}
col1 = "#011f4b"
brooklyn + geom_point(data = databrooklyn, aes(x = longitude, y = latitude), alpha=0.002, color=col1) + geom_jitter()
ggsave("Brooklyn_accidents.png", dpi=300)
```



```{r}

#frequency table for contributing factor 1
brookcf1 <- as.data.frame(table(databrooklyn$contributing_factor_vehicle_1))
colnames(brookcf1) <- c("cf1","freq")
brookcf1 <- brookcf1 %>% arrange(desc(freq))

#function for creating frequency tables:
func_tabcf <- function(colname, newname){
  temp <- as.data.frame(table(databrooklyn[,colname]))
  colnames(temp) <- c(newname, "freq")
  return(temp%>%arrange(desc(freq)))
}

#frequency table for contributing factor 2
brookcf2 <- func_tabcf("contributing_factor_vehicle_2","cf2")
#frequency table for contributing factor 3
brookcf3 <- func_tabcf("contributing_factor_vehicle_3","cf3")
#frequency table for contributing factor 4
brookcf4 <- func_tabcf("contributing_factor_vehicle_4","cf4")
#frequency table for contributing factor 5
brookcf5 <- func_tabcf("contributing_factor_vehicle_5","cf5")


#processing vehicle code:
#converting vehicle code to upper case:
databrooklyn$vehicle_type_code1 <- toupper(databrooklyn$vehicle_type_code1)
databrooklyn$vehicle_type_code2 <- toupper(databrooklyn$vehicle_type_code2)
databrooklyn$vehicle_type_code_3 <- toupper(databrooklyn$vehicle_type_code_3)
databrooklyn$vehicle_type_code_4 <- toupper(databrooklyn$vehicle_type_code_4)
databrooklyn$vehicle_type_code_5 <- toupper(databrooklyn$vehicle_type_code_5)

#creating frequency tables for vehicle code from 1 to 5
brookvc1 <- func_tabcf("vehicle_type_code1","vc1")
brookvc2 <- func_tabcf("vehicle_type_code2","vc2")
brookvc3 <- func_tabcf("vehicle_type_code_3","vc3")
brookvc4 <- func_tabcf("vehicle_type_code_4","vc4")
brookvc5 <- func_tabcf("vehicle_type_code_5","vc5")

#creating frequency table for zipcode:
brookzc <- func_tabcf("zip_code","zipcode")
```

```{r}
## adding lat/lon to zipcodes using geocode api:
brookzc$zipcode <- as.character(brookzc$zipcode)
brookzc <- cbind(brookzc, geocode(brookzc$zipcode))
```

## overlaying top ten areas by zipcode where Brooklyn has accidents:
```{r}
brookzc$zipcode <- as.factor(brookzc$zipcode)

brooklyn + geom_point(data = databrooklyn, aes(x = longitude, y = latitude, size=total_injured), alpha=0.002, color="blue") +
  geom_point(aes(x = lon, y = lat, shape=zipcode, stroke = 2,size=2.5), colour="red", data = brookzc[1:10,]) + 
  scale_shape_manual(values=1:nlevels(brookzc$zipcode))
```

```{r}
#adding label/rank to zipcode
brookzc$label <- paste(row.names(brookzc),brookzc$zipcode,sep=', ')
colnames(brookzc)[2] <- "total accidents"

brooklyn + geom_point(data = databrooklyn, aes(x = longitude, y = latitude), alpha=0.002, color="blue") +
  geom_point(aes(x = lon, y = lat, stroke = 2,size=`total accidents`), colour="red", data = brookzc[1:15,]) + 
   geom_label_repel(
    aes(lon, lat, label = label),
    data=brookzc[1:15,],
    family = 'Arial', 
    size = 3, 
    box.padding = 0.2, point.padding = 0.3,
    segment.color = 'black')  + labs(size = "Accidents In Major Areas") +xlab("Longitude") + ylab("Latitude")
ggsave("Brook_top10accbyzc.png", dpi=300) 
```

### histogram of zipcode vs number of accidents:
```{r}
ggplot(data=brookzc[1:15,], aes(x=reorder(zipcode,-`total accidents`),y=`total accidents`))+geom_bar(stat="identity",color="blue", fill="blue")+scale_y_continuous(breaks=c(10000,15000,20000),labels=c('10K','15K','20K'))+xlab("Zipcode")+ylab("Number of Accidents")+theme_light()
```

### histogram of hour of the day vs number of accidents
```{r}
ggplot(data=databrooklyn, aes(x=hour))+geom_histogram(stat="count",color="blue", fill="blue")+xlab("Hour of the Day")+ylab("Number of Accidents")+scale_x_continuous(breaks=seq(0,23,1))+theme_light()
```

```{r}
#creating frequency table of hour
brookhour <- func_tabcf("hour","hour")
```

### histogram of year vs number of accidents
```{r}
ggplot(data=databrooklyn, aes(x=year))+geom_histogram(stat="count",color="blue", fill="blue")+xlab("Year")+ylab("Number of Accidents")+scale_x_continuous(breaks=seq(2012,2019,1))+theme_light()
```

### histogram of month vs number of accidents
```{r}
ggplot(data=databrooklyn, aes(x=month))+geom_histogram(stat="count",color="blue", fill="blue")+xlab("Month")+ylab("Number of Accidents")+scale_x_continuous(breaks=seq(1,12,1))+theme_light()
```

### percentage of accidents during top 10 peak hours:
```{r}
sum(brookhour$freq[1:11])/sum(brookhour$freq);
sum(brookhour$freq[1:11]);
```

### filtering brooklyn data by peak hours of accidents from 8 am to 6 pm:
```{r}
dat_brookpkhr <- subset(databrooklyn, databrooklyn$hour>=8 & databrooklyn$hour<=18)
brookzcpkhr <- as.data.frame(table(dat_brookpkhr$zip_code))
colnames(brookzcpkhr)[1] <- "zcpkhr"
brookzcpkhr <- brookzcpkhr %>% arrange(desc(Freq))
```
```{r}
zctop15pkhr <- brookzcpkhr$zcpkhr[1:15]
zctop15all <- brookzc$zipcode[1:15]
#note: zctop15pkhr and zctop15all are 93.33% same
```

```{r}
## checking top 15 areas of zipcodes over total accidents during peak hours:
sum(brookzcpkhr$`total accidents`[1:15])/sum(brookzcpkhr$`total accidents`);
sum(brookzcpkhr$`total accidents`[1:15])
sum(brookzcpkhr$`total accidents`);
```

##overlaying top 15 areas during peak hours by zipcode where Brooklyn has accidents:
```{r}
## adding lat/lon to zipcodes pk hour:
brookzcpkhr$zcpkhr <- as.character(brookzcpkhr$zcpkhr)
brookzcpkhr <- cbind(brookzcpkhr, geocode(brookzcpkhr$zcpkhr))

#adding label/rank to zipcode pk hour
brookzcpkhr$label <- paste(row.names(brookzcpkhr),brookzcpkhr$zcpkhr,sep=', ')
colnames(brookzcpkhr)[2] <- "total accidents"

brooklyn + geom_point(data = databrooklyn, aes(x = longitude, y = latitude), alpha=0.002, color="blue") +
  geom_point(aes(x = lon, y = lat, stroke = 2,size=`total accidents`), colour="red", data = brookzcpkhr[1:15,]) + 
   geom_label_repel(
    aes(lon, lat, label = label),
    data=brookzcpkhr[1:15,],
    family = 'Calibri', 
    size = 3, 
    box.padding = 0.2, point.padding = 0.3,
    segment.color = 'black')  + labs(size = "Accidents In Major Areas") +xlab("Longitude") + ylab("Latitude")
ggsave("Brook_top15byzcpkhr.png", dpi=300) 
```

## Percentage of Accident Happen in Top 15 Areas during Peak Hour:
```{r}
sum(brookzcpkhr$`total accidents`[1:15])/sum(brookzcpkhr$`total accidents`);
sum(brookzcpkhr$`total accidents`[1:15]);
sum(brookzcpkhr$`total accidents`);
```

## Analyzing factors which are main causes of accidents:
```{r}
func_freq_desc <- function(data,colname,newname){
  temp <- as.data.frame(table(data[,colname]))
  colnames(temp) <- c(newname, "freq")
  return(temp%>%arrange(desc(freq)))
} 

brookcf1pkhr <- func_freq_desc(dat_brookpkhr,"contributing_factor_vehicle_1","cf1pk")
brookcf2pkhr <- func_freq_desc(dat_brookpkhr,"contributing_factor_vehicle_2","cf2pk")
brookcf3pkhr <- func_freq_desc(dat_brookpkhr,"contributing_factor_vehicle_3","cf3pk")
brookcf4pkhr <- func_freq_desc(dat_brookpkhr,"contributing_factor_vehicle_4","cf4pk")
brookcf5pkhr <- func_freq_desc(dat_brookpkhr,"contributing_factor_vehicle_5","cf5pk")
#major causes of accidents for vehicle one
sum(brookcf1pkhr$freq[2:21])/sum(brookcf1pkhr$freq)
```


## Grouping causes/factors of accidents into 8(including unspecified) main categories
- Driving Mistakes
- Vehicle Problems
- Route Defects
- Health
- Breaking Rules
- Other
- Distraction
- Unspecified

```{r}
## loading data which contains this factors categories:
cf6cat <- xlsx::read.xlsx(file = "data/eda.xlsx", sheetIndex = 1, encoding = "UTF-8")
datamaster <- databrooklyn
colnames(cf6cat)[1] <- "contributing_factor_vehicle_1"
cf6cat$contributing_factor_vehicle_1 <- as.character(cf6cat$contributing_factor_vehicle_1)
databrooklynnew <- join_all(list(databrooklyn,cf6cat),by="contributing_factor_vehicle_1",type="left")
table(databrooklynnew$CFType)
```

```{r}
### checking why it was not merging properly
if (str_sub(databrooklynnew$contributing_factor_vehicle_1[16],28,n)==str_sub(cf6cat$contributing_factor_vehicle_1[46],28,n)){print(str_sub(databrooklynnew$contributing_factor_vehicle_1[16],28,n));
  print((databrooklynnew$contributing_factor_vehicle_1[16]))
  print(cf6cat$contributing_factor_vehicle_1[46]); print("true")}
```

```{r}
## Removing white spaces in contributing factor
databrooklyn$cfv1 <- str_replace_all(databrooklyn$contributing_factor_vehicle_1,"\\s","")
databrooklyn$cfv2 <- str_replace_all(databrooklyn$contributing_factor_vehicle_2,"\\s","")
databrooklyn$cfv3 <- str_replace_all(databrooklyn$contributing_factor_vehicle_3,"\\s","")
databrooklyn$cfv4 <- str_replace_all(databrooklyn$contributing_factor_vehicle_4,"\\s","")
databrooklyn$cfv5 <- str_replace_all(databrooklyn$contributing_factor_vehicle_5,"\\s","")
```

```{r}
## loading again:
databrooklynnew <-NULL
## processing for contributing factor 1:
cf6cat$cfv1 <- str_replace_all(cf6cat$contributing_factor_vehicle_1,"\\s","")
colnames(cf6cat)[2] <- "cftype1"
databrooklynnew <- join_all(list(databrooklyn,cf6cat),by="cfv1",type="left")
table(databrooklynnew$cftype1,useNA="always")
```


```{r}
#processing for contributing factor 2
cf6cat$cfv2 <- str_replace_all(cf6cat$contributing_factor_vehicle_1,"\\s","")
colnames(cf6cat)[2]<- "cftype2"
databrooklynnew <- join_all(list(databrooklynnew,cf6cat),by="cfv2",type="left")
table(databrooklynnew$cftype2,useNA="always")
```

```{r}
#processing for contributing factor 3
cf6cat$cfv3 <- str_replace_all(cf6cat$contributing_factor_vehicle_1,"\\s","")
colnames(cf6cat)[2]<- "cftype3"
databrooklynnew <- join_all(list(databrooklynnew,cf6cat),by="cfv3",type="left")
table(databrooklynnew$cftype3,useNA="always")
```


```{r}
#processing for contributing factor 4
cf6cat$cfv4 <- str_replace_all(cf6cat$contributing_factor_vehicle_1,"\\s","")
colnames(cf6cat)[2]<- "cftype4"
databrooklynnew <- join_all(list(databrooklynnew,cf6cat),by="cfv4",type="left")
table(databrooklynnew$cftype4,useNA="always")
```


```{r}
#processing for contributing factor 5
cf6cat$cfv5 <- str_replace_all(cf6cat$contributing_factor_vehicle_1,"\\s","")
colnames(cf6cat)[2]<- "cftype5"
databrooklynnew <- join_all(list(databrooklynnew,cf6cat),by="cfv5",type="left")
table(databrooklynnew$cftype5,useNA="always")
```

```{r}
## subsetting the columns which are required:
databrooklynnew <- databrooklynnew[,c("unique_key","cftype1","cftype2","cftype3","cftype4","cftype5")]
```

```{r}
## merging it with databrooklyn:
databrooklynnew <- join_all(list(databrooklyn,databrooklynnew),by="unique_key",type="inner")
```

### Histogram of cftype1 vs number of accidents for all Brooklyn
```{r}
ggplot(data=subset(databrooklynnew,!is.na(databrooklynnew$cftype1)&databrooklynnew$cftype1!="Unspecified"), aes(x=cftype1))+geom_bar(stat="count",color="blue", fill="blue")+xlab("Vehicle 1 Contributing Factor ")+ylab("Number of Accidents")+theme(axis.text.x=element_text(family = "Times",angle=30,hjust=1,vjust=1))+
    geom_text(
   stat='count', aes(label=paste0(round(..count../sum(..count..)*100),'%')), vjust=-0.1
  )+ggtitle("                Number of Accidents by Vehicle 1 Contributing Factor ")
ggsave("cftype1.png")
```

```{r}
nrow(subset(databrooklynnew,!is.na(databrooklynnew$cftype1)&databrooklynnew$cftype1!="Unspecified"))
```

### Histogram of cftype2 vs number of accidents for all Brooklyn
```{r}
ggplot(data=subset(databrooklynnew,!is.na(databrooklynnew$cftype2)&databrooklynnew$cftype2!="Unspecified"), aes(x=cftype2))+geom_bar(stat="count",color="blue", fill="blue")+xlab("Vehicle 2 Contributing Factor ")+ylab("Number of Accidents")+theme(axis.text.x=element_text(family = "Times",angle=30,hjust=1,vjust=1))+
    geom_text(
   stat='count', aes(label=paste0(round(..count../sum(..count..)*100),'%')), vjust=-0.1
  )+ggtitle("                Number of Accidents by Vehicle 2 Contributing Factor ")
ggsave("cftype2.png")
```


#Histogram of cftype3 vs number of accidents for all Brooklyn
```{r}
ggplot(data=subset(databrooklynnew,!is.na(databrooklynnew$cftype3)&databrooklynnew$cftype3!="Unspecified"), aes(x=cftype3))+geom_bar(stat="count",color="blue", fill="blue")+xlab("Vehicle 3 Contributing Factor ")+ylab("Number of Accidents")+theme(axis.text.x=element_text(family = "Times",angle=30,hjust=1,vjust=1))+
    geom_text(
   stat='count', aes(label=paste0(round(..count../sum(..count..)*100),'%')), vjust=-0.1
  )+ggtitle("                Number of Accidents by Vehicle 3 Contributing Factor ")
ggsave("cftype3.png")
```


```{r}
## grouping by cftype 1 and cftype2 separately:
databrooklynnew%>%dplyr::group_by(cftype1)%>%tally();
databrooklynnew%>%dplyr::group_by(cftype2)%>%tally();
```

```{r}
databrooklynnew%>%dplyr::group_by(cftype1,cftype2)%>%tally()
```

### looking at the relationship between cftype1 and cftype2:
```{r}
top81 <- c('Breaking Rules','Distraction/Inattention','Driving Mistakes','Health','Route Defects','Vehicle Problems','Other')
subset(databrooklynnew, databrooklynnew$cftype1 %in% top81 & databrooklynnew$cftype2 %in% top81) %>% 
  select(cftype1,cftype2,unique_key) %>% 
  dplyr::group_by(cftype1,cftype2) %>% 
  dplyr::summarise(total=n())%>%
  spread(cftype2,total)
```

```{r}
cftype1 <- func_freq_desc(databrooklynnew,"cftype1","cftype1")
#add percentage column
cftype1 <- cftype1%>%mutate(percentall = freq/sum(freq)*100)
```

### Contributing Factor Type Association 1 with Injury/Deaths for overall
```{r}
graphdicf1 <- databrooklynnew[(!is.na(databrooklynnew$cftype1) & databrooklynnew$cftype1!="Unspecified"),]%>%group_by(cftype1)%>%
  dplyr::summarise(total_injured=sum(total_injured,na.rm = T),total_killed=sum(total_killed,na.rm = T))%>%  
  dplyr::arrange(desc(total_injured))%>%
  mutate(percent_injured=total_injured/sum(total_injured)*100)%>%
  mutate(percent_killed=total_killed/sum(total_killed)*100)
```


```{r}

scaleFactor <- max(graphdicf1$total_injured) / max(graphdicf1$total_killed)
plotdicf1 <- ggplot(graphdicf1, aes(x=cftype1,  width=.4)) +
  geom_col(aes(y=total_injured), fill="blue", position = position_nudge(x = -.4)) +
  geom_text(
   stat='identity', aes(label=paste0(round(percent_injured),"%"),y = total_injured), vjust=-0.1,nudge_x = -0.38
  )+
  geom_col(aes(y=total_killed*scaleFactor), fill="red") +
  geom_text(
   stat='identity', aes(label=paste0(round(percent_killed),"%"),y = total_killed*scaleFactor), vjust=-0.1,nudge_x = 0.1
  )+
  scale_y_continuous(name="Total Injured", sec.axis=sec_axis(~./scaleFactor, name="Total Killed")) +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red"),axis.text.x=element_text(family = "Times",angle=30,hjust=1,vjust=1)
  ) + 
  labs(title = "       Total Persons Injured and Killed vs Contributing Factor Vechicle 1")+xlab("Vehicle 1 Contributing Factor")
plotdicf1
#ggsave("deathinjuredcf1.png")
```


### Contributing Factor Type Association 2 with Injury/Deaths for overall
```{r}
graphdicf2 <- databrooklynnew[(!is.na(databrooklynnew$cftype2) & databrooklynnew$cftype2!="Unspecified"),]%>%group_by(cftype2)%>%
  dplyr::summarise(total_injured=sum(total_injured,na.rm = T),total_killed=sum(total_killed,na.rm = T))%>%  
  dplyr::arrange(desc(total_injured))%>%
  mutate(percent_injured=total_injured/sum(total_injured)*100)%>%
  mutate(percent_killed=total_killed/sum(total_killed)*100)
```

```{r}
scaleFactor <- max(graphdicf2$total_injured) / max(graphdicf1$total_killed)
ggplot(graphdicf2, aes(x=cftype2,  width=.4)) +
  geom_col(aes(y=total_injured), fill="blue", position = position_nudge(x = -.4)) +
  geom_text(
   stat='identity', aes(label=paste0(round(percent_injured),"%"),y = total_injured), vjust=-0.1,nudge_x = -0.38
  )+
  geom_col(aes(y=total_killed*scaleFactor), fill="red") +
  geom_text(
   stat='identity', aes(label=paste0(round(percent_killed),"%"),y = total_killed*scaleFactor), vjust=-0.1,nudge_x = 0.1
  )+
  scale_y_continuous(name="Total Injured", sec.axis=sec_axis(~./scaleFactor, name="Total Killed")) +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red"),axis.text.x=element_text(family = "Times",angle=30,hjust=1,vjust=1)
  ) + 
  labs(title = "       Total Persons Injured and Killed vs Contributing Factor Vechicle 2")+xlab("Vehicle 2 Contributing Factor")
#ggsave("deathinjuredcf2.png")

```


```{r}
##replacing same car type different names to same names
databrooklynnew$vehicle_type_code1 <- str_replace(databrooklynnew$vehicle_type_code1,"SPORT UTILITY / STATION WAGON","SPORTS CAR/STATION WAGON")
databrooklynnew$vehicle_type_code1 <- str_replace(databrooklynnew$vehicle_type_code1,"STATION WAGON/SPORT UTILITY VEHICLE","SPORTS CAR/STATION WAGON")
databrooklynnew$vehicle_type_code2 <- str_replace(databrooklynnew$vehicle_type_code2,"SPORT UTILITY / STATION WAGON","SPORTS CAR/STATION WAGON")
databrooklynnew$vehicle_type_code2 <- str_replace(databrooklynnew$vehicle_type_code2,"STATION WAGON/SPORT UTILITY VEHICLE","SPORTS CAR/STATION WAGON")
```

### subsetting databrooklyn containing only those vehicles which have top 15 vehicles:
```{r}
### Note: Top 15 vehicles in peak hour are same as overall 
brookvc1 <- func_freq_desc(databrooklynnew,"vehicle_type_code1","vc1")
```


```{r}
## Note from above: Sports Utility Vehcicle are the second highest to have distractions, driving mistakes, breaking rules, other, route defects, unspecified:
#SPORT UTILITY / STATION WAGON same as below                 
# STATION WAGON/SPORT UTILITY VEHICLE
```

```{r}
#Top 5 percentage in overall excluding other types:
(sum(brookvc1$freq[1:6])-6784)/(sum(brookvc1$freq))
```

### Subsetting brookylndata containing top 5 vehicle code1:
```{r}
top5vc1 <- brookvc1$vc1[1:6]
top5vc1 <- top5vc1[-4]
databrooktop5vc1 <- subset(databrooklynnew, databrooklynnew$vehicle_type_code1 %in% top5vc1)
table(databrooktop5vc1$vehicle_type_code1,databrooktop5vc1$cftype1)

```

### Relationship between top5vc1 and cftype1:
```{r}
summarycf1vc1 <- databrooktop5vc1[(!is.na(databrooktop5vc1$cftype1) &       databrooktop5vc1$cftype1!="Unspecified"),]%>%group_by(cftype1,vehicle_type_code1)%>%
  dplyr::summarise(n=n())%>%spread(cftype1,value = n)%>%arrange(desc(`Breaking Rules`))%>%
  mutate(BRpercent=round(`Breaking Rules`/sum(`Breaking Rules`)*100,2))%>%
  mutate(DIpercent=round(`Distraction/Inattention`/sum(`Distraction/Inattention`)*100,2))%>%
  mutate(DMpercent=round(`Driving Mistakes`/sum(`Driving Mistakes`)*100,2))%>%
  mutate(Hpercent=round(`Health`/sum(`Health`)*100,2))%>%
  mutate(Opercent=round(`Other`/sum(`Other`)*100,2))%>%
  mutate(RDpercent=round(`Route Defects`/sum(`Route Defects`)*100,2))%>%
  mutate(VPpercent=round(`Vehicle Problems`/sum(`Vehicle Problems`)*100,2))

  summarycf1vc1[,2:ncol(summarycf1vc1)] = apply(summarycf1vc1[,2:ncol(summarycf1vc1)], 2, function(x) as.character(x))
  for ( i in 1:nrow(summarycf1vc1)) {
    
    for ( j in 2:8){
     summarycf1vc1[i,j] <- paste(summarycf1vc1[i,j]," (",trimws(summarycf1vc1[i,7+j],"both"),"%)",sep = "") 
    }
    }
  summarycf1vc1[,c(9:ncol(summarycf1vc1))] <- NULL
```

### vehicle 1 vs contributing factor type 2:
```{r}
table(databrooktop5vc1$vehicle_type_code1,databrooktop5vc1$cftype2)
```

### vehicle code 1 Association with Injury/Deaths for overall:
```{r}
databrooktop5vc1[!is.na(databrooktop5vc1$vehicle_type_code1),]%>%group_by(vehicle_type_code1)%>%
  dplyr::summarise(total_injured=sum(total_injured,na.rm = T),total_killed=sum(total_killed,na.rm = T))%>%  
  dplyr::arrange(desc(total_injured))%>%
  mutate(percent_injured=total_injured/sum(total_injured)*100)%>%
  mutate(percent_killed=total_killed/sum(total_killed)*100)
```

### vehcicle code 2 Association with Injury/Deaths for overall:
```{r}
databrooktop5vc1[!is.na(databrooktop5vc1$vehicle_type_code2),]%>%group_by(vehicle_type_code2)%>%
  dplyr::summarise(total_injured=sum(total_injured,na.rm = T),total_killed=sum(total_killed,na.rm = T))%>%  
  dplyr::arrange(desc(total_injured))%>%
  mutate(percent_injured=total_injured/sum(total_injured)*100)%>%
  mutate(percent_killed=total_killed/sum(total_killed)*100)
```

### Relationship between vehicle code 2 and vehicle code 1 for data subset: databrooklyntop5vc1:
```{r}
table(databrooktop5vc2$vehicle_type_code1,databrooktop5vc2$vehicle_type_code2)
```

```{r}
### Probability of contributing factor type 1 having an accident with another vehicle (vehicle code 2)
table(databrooktop5vc1$cftype1,databrooktop5vc1$vehicle_type_code2)
```

### Subsetting brookylndata containing top 5 vehicle code2:
```{r}
brookvc2 <- func_freq_desc(databrooklynnew, "vehicle_type_code2" , "vc2")
top5vc2 <- brookvc2$vc2[1:7];
top5vc2 <- top5vc2[-c(3,6)];
top5vc2;
databrooktop5vc2 <- subset(databrooklynnew, databrooklynnew$vehicle_type_code2 %in% top5vc2)
```

### Relationship between top5vc2 vs cftype1
```{r}
summarycf1vc2 <- databrooktop5vc2[(!is.na(databrooktop5vc2$cftype1) &       databrooktop5vc2$cftype1!="Unspecified"),]%>%group_by(cftype1,vehicle_type_code2)%>%
  dplyr::summarise(n=n())%>%spread(cftype1,value = n)%>%arrange(desc(`Breaking Rules`))%>%
  mutate(BRpercent=round(`Breaking Rules`/sum(`Breaking Rules`)*100,2))%>%
  mutate(DIpercent=round(`Distraction/Inattention`/sum(`Distraction/Inattention`)*100,2))%>%
  mutate(DMpercent=round(`Driving Mistakes`/sum(`Driving Mistakes`)*100,2))%>%
  mutate(Hpercent=round(`Health`/sum(`Health`)*100,2))%>%
  mutate(Opercent=round(`Other`/sum(`Other`)*100,2))%>%
  mutate(RDpercent=round(`Route Defects`/sum(`Route Defects`)*100,2))%>%
  mutate(VPpercent=round(`Vehicle Problems`/sum(`Vehicle Problems`)*100,2))

 summarycf1vc2[,2:ncol(summarycf1vc2)] = apply(summarycf1vc2[,2:ncol(summarycf1vc2)], 2, function(x) as.character(x))

  for ( i in 1:nrow(summarycf1vc2)) {
    
    for ( j in 2:8){
      
     summarycf1vc2[i,j] <- paste(summarycf1vc2[i,j]," (",trimws(summarycf1vc2[i,7+j],"both"),"%)",sep = "") 
    }
  }
 
  summarycf1vc2[,c(9:ncol(summarycf1vc2))] <- NULL
  
```

### Subsetting data containing top 5 vehicle code for 1 and 2
```{r}
databrooktop5vc1vc2 <- subset(databrooklynnew, databrooklynnew$vehicle_type_code1 %in% top5vc1 & databrooklynnew$vehicle_type_code2 %in% top5vc2)
```

### relationship between vehicle code 1 and vehicle code 2
```{r}
summaryvc1vc2 <- databrooktop5vc1vc2%>%dplyr::group_by(vehicle_type_code1,vehicle_type_code2)%>%
  dplyr::summarise(n=n())%>%
  spread(vehicle_type_code1,n)%>%
  arrange(desc(`SPORTS CAR/STATION WAGON`))%>%
  mutate(PVperent=round(`PASSENGER VEHICLE`/sum(`PASSENGER VEHICLE`,na.rm = T)*100,2))%>%
  mutate(PUTpercent=round(`PICK-UP TRUCK`/sum(`PICK-UP TRUCK`,na.rm = T)*100,2))%>%
  mutate(Spercent=round(`SEDAN`/sum(`SEDAN`,na.rm = T)*100,2))%>%
  mutate(Sportpercent=round(`SPORTS CAR/STATION WAGON`/sum(`SPORTS CAR/STATION WAGON`,na.rm = T)*100,2))%>%
  mutate(Vpercent=round(`VAN`/sum(`VAN`,na.rm = T)*100,2))
  
 summaryvc1vc2[,2:ncol(summaryvc1vc2)] = apply(summaryvc1vc2[,2:ncol(summaryvc1vc2)], 2, function(x) as.character(x))

  for ( i in 1:nrow(summaryvc1vc2)) {
    for ( j in 2:6){
     summaryvc1vc2[i,j] <- paste(summaryvc1vc2[i,j]," (",trimws(summaryvc1vc2[i,5+j],"both"),"%)",sep = "") 
    }
  }
  summaryvc1vc2[,c(7:ncol(summaryvc1vc2))] <- NULL
  summaryvc1vc2 <- summaryvc1vc2[,c(1,5,2,4,6,3)]
```


### grouping top5vc1 by vc1 and summarizing total injured and total killed:
```{r}
databrooktop5vc1%>%dplyr::group_by(vehicle_type_code1)%>%
  dplyr::summarise(total_injured=sum(total_injured,na.rm = T),total_killed=sum(total_killed,na.rm = T))%>%  
  dplyr::arrange(desc(total_injured))%>%
  mutate(percent_injured=total_injured/sum(total_injured)*100)%>%
  mutate(percent_killed=total_killed/sum(total_killed)*100)
```

### grouping top5vc2 by vc2 and summarizing total injured and total killed:
```{r}
databrooktop5vc2%>%dplyr::group_by(vehicle_type_code2)%>%
  dplyr::summarise(total_injured=sum(total_injured,na.rm = T),total_killed=sum(total_killed,na.rm = T))%>%  
  dplyr::arrange(desc(total_injured))%>%
  mutate(percent_injured=total_injured/sum(total_injured)*100)%>%
  mutate(percent_killed=total_killed/sum(total_killed)*100)
```

### grouping top5vc1vc2 by vc1 and vc2 and summarizing total injured and total killed:
### Basically pairing top5 vehicle code 1 with top 5 vehicle code 2:
```{r}
summarydivc1vc2 <- databrooktop5vc1vc2%>%dplyr::group_by(vehicle_type_code1,vehicle_type_code2)%>%
  dplyr::summarise(total_injured=sum(total_injured,na.rm = T),total_killed=sum(total_killed,na.rm = T))%>%  
  dplyr::arrange(desc(total_injured),desc(total_killed))%>%
  mutate(percent_injured=round(total_injured/sum(total_injured)*100,2))%>%
  mutate(percent_killed=round(total_killed/sum(total_killed)*100,2))

 summarydivc1vc2[,3:ncol(summarydivc1vc2)] = apply(summarydivc1vc2[,3:ncol(summarydivc1vc2)], 2, function(x) as.character(x))

  for ( i in 1:nrow(summarydivc1vc2)) {
    for ( j in 3:4){
     summarydivc1vc2[i,j] <- paste(summarydivc1vc2[i,j]," (",trimws(summarydivc1vc2[i,2+j],"both"),"%)",sep = "") 
    }
  }
 summarydivc1vc2[,c(5:ncol(summarydivc1vc2))] <- NULL
```

### subsettting top 10 pariring of vc1 and vc2:
```{r}
summarydivc1vc2 <- summarydivc1vc2[1:10,]
summarydivc1vc2$vehicle_type_code1 <- str_replace(summarydivc1vc2$vehicle_type_code1,"PASSENGER VEHICLE","PV1")
summarydivc1vc2$vehicle_type_code2 <- str_replace(summarydivc1vc2$vehicle_type_code2,"PASSENGER VEHICLE","PV2")
summarydivc1vc2$vehicle_type_code1 <- str_replace(summarydivc1vc2$vehicle_type_code1,"SPORTS CAR/STATION WAGON","SCSW1")
summarydivc1vc2$vehicle_type_code2 <- str_replace(summarydivc1vc2$vehicle_type_code2,"SPORTS CAR/STATION WAGON","SCSW2")
summarydivc1vc2$vehicle_type_code1 <- str_replace(summarydivc1vc2$vehicle_type_code1,"SEDAN","SEDAN1")
summarydivc1vc2$vehicle_type_code2 <- str_replace(summarydivc1vc2$vehicle_type_code2,"SEDAN","SEDAN2")
summarydivc1vc2$vehicle_type_code2 <- str_replace(summarydivc1vc2$vehicle_type_code2,"BICYCLE","BICYCLE2")
summarydivc1vc2$vehicle_type_code2 <- str_replace(summarydivc1vc2$vehicle_type_code2,"VAN","VAN2")
```


### Histogram of pairing of vc1 and vc2, showing total injured and total killed:
```{r}
summarydivc1vc2$vc1vc2 <- paste0(summarydivc1vc2$vehicle_type_code1,"-",summarydivc1vc2$vehicle_type_code2)
scaleFactor <- max(summarydivc1vc2$total_injured) / max(summarydivc1vc2$total_killed)
ggplot(summarydivc1vc2, aes(x=vc1vc2,  width=.4)) +
  geom_col(aes(y=total_injured), fill="blue", position = position_nudge(x = -.4)) +
  geom_text(
   stat='identity', aes(label=paste0(round(percent_injured),"%"),y = total_injured), vjust=-0.1,nudge_x = -0.38
  )+
  geom_col(aes(y=total_killed*scaleFactor), fill="red") +
  geom_text(
   stat='identity', aes(label=paste0(round(percent_killed),"%"),y = total_killed*scaleFactor), vjust=-0.1,nudge_x = 0.1
  )+
  scale_y_continuous(name="Total Injured", sec.axis=sec_axis(~./scaleFactor, name="Total Killed")) +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red"),axis.text.x=element_text(family = "Times",angle=30,hjust=1,vjust=1)
  ) + 
  labs(title = "Total Persons Injured and Killed when Vehicle Type1 hits another Vehicle Type 2")+xlab("Vehicle Type 1 - Vehicle Type 2")
ggsave("deathinjuredvc1vc2.png")
```



